---
title: "p8105 Homework 2"
author: Rebekah Hughes
output: github_document
---

In the code chunk below, all the packages used in this homework are loaded.

```{r}
library(tidyverse)
library(readxl)
```


## Problem 1

The following code reads and cleans the Mr. Trash Wheel dataset.

```{r}
mrtrash_df = 
  read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", 
  sheet = "Mr. Trash Wheel",
  range = "A2:N408") %>%
  janitor::clean_names() %>%
  drop_na (dumpster) %>% 
  mutate (round (sports_balls, digits = 0), as.integer(sports_balls))
```

The following code reads and cleans the Precipitation data for 2017 and 2018 and combines the datasets into one.

```{r}
precip_2017 =
  read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
             sheet = "2017 Precipitation",
             range = "A2:B14") %>% 
  mutate(precip_year = 2017)

precip_2018 =
  read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
             sheet = "2018 Precipitation",
             range = "A2:B14") %>% 
  mutate(precip_year = 2018)

precip_tidy =
  bind_rows(precip_2017, precip_2018) %>% 
  janitor::clean_names() %>% 
  mutate (month.name [month])
```

The above datasets used for problem 1 have the following numbers of observations: `r nrow(mrtrash_df)` for the Mr. Trash Wheel dataset; `r nrow(precip_2017)` for the 2017 Precipitation dataset; `r nrow(precip_2018)` for the 2018 Precipitation dataset; and `r nrow(precip_tidy)` for the combined precipitation dataset. The following are examples of key variables from the Mr. Trash Wheel dataset: the mean of weight is `r mean (mrtrash_df$weight_tons)` tons, with a maximum value of `r max(mrtrash_df$weight_tons)` tons and a minimum value of `r min(mrtrash_df$weight_tons)` tons; the mean amount of grocery bags is `r mean(mrtrash_df$grocery_bags)`; and the mean amount of cigarette butts are `r mean(mrtrash_df$cigarette_butts)`, with a maximum of `r max(mrtrash_df$cigarette_butts)` and a minimum of `r min(mrtrash_df$cigarette_butts)`. The following are examples of key variables from the combined precipitation dataset and the two individual precipitation datasets: the mean total rainfall across the two years is `r mean(precip_tidy$total)` inches; the mean total rainfall for 2017 alone is `r mean(precip_2017$Total)` inches; and the mean total rainfall for 2018 is `r mean(precip_2018$Total)` inches. The total precipitation in 2018 was `r sum (precip_2018$Total)` and the median number of sports balls in a dumpster in 2017 was `r median (mrtrash_df$sports_balls, "year" == 2017)`.

## Problem 2

```{r}
nycsub_df = 
  read_csv ("./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>%
  janitor::clean_names() %>%
  select(line, station_name, station_latitude, station_longitude, route1, route2, route3, route4, route5, route6, route7, route8, route9, route10, route11, entry, vending, entrance_type, ada) %>%
  mutate(recode(entry, YES = TRUE, NO = FALSE))
```

The above dataset contains the line, station name, routes 1-11, entry, vending and entrance type as character variables, station latitude and longitude as nunerical variables, and ADA compliance and the recoded entry variable as logical variables. So far to clean the data, clean names has been done, certain variables have been cut out of the dataset, and entry has been recoded as a logical variable. The dimensions of the dataset are `r nrow(nycsub_df)` rows by `r ncol(nycsub_df)` columns. The data is more tidy than what it was initally but could be made more tidy still.



